---
title: "Fitness_tracker_project1"
author: "Ryan King"
date: "12/4/2021"
output: html_document
---
### Project 1: Fitness tracker

## Load the data and environment
```{r file_import}
library(dplyr)
library(ggplot2)
library(data.table)
file_path <- "C:/Users/I0485672/Downloads/activity.csv"
data <- read.csv(file_path)
```

## Analyzing the data

There are some NA's in here. They will be removed until imputed step.
Why? They don't have the device on or working.
We cannot say that means zero steps.

# Let's look at the total number of steps taken each day
```{r daily_sum_of_steps}
data_summary <- data[complete.cases(data), ] %>% group_by(date) %>% summarize(mean=sum(steps, na.rm = TRUE))
data_summary

data_summary %>% ggplot(aes(x = mean)) + geom_histogram(bins = 20) +
  theme_bw() + labs(x='Mean steps per day', y='Number of occurances', 
                    title = 'Frequency of mean steps per day')
```

# let's look at the mean and steps per day.
```{r mean_and_median}
mean(data_summary$StepsSum)
median(data_summary$StepsSum)
```

# I see more zeros than numbers, which means median will be 0. Is this true?
table(data %>% filter(date == '2012-10-03')  %>% .[[1]] > 0)
```
It's a class imbalance... more zeros than numbers, so it's correct

## let's look at a time series
```{r timeseries_mean}
data_summary <-  data[complete.cases(data), ] %>% group_by(interval) %>% summarize(mean=mean(steps, na.rm = TRUE))

ggplot(data_summary, aes(x = interval, y = mean)) + 
  geom_point() + geom_line() + theme_bw() + 
  labs(y="Mean steps", x = "Minutes from start of day", 
       title = "Average number of steps during the day")
```

let's look at the highest steps taken
```{r highest steps}
data_summary <-  data[complete.cases(data), ] %>% group_by(interval) %>% summarize(mean=mean(steps, na.rm = TRUE))
data_summary %>% arrange(desc(mean))

print("The largest interval")
data_summary %>% arrange(desc(mean)) %>% head(n=1)
```

## Imputing data

How to impute data? There's many many theories out there
1. mean - on each interval
2. nearest neighbor - for example, steps on a Saturday vs. Monday and take mean
3. So much more, but #2 is really good.

Let's do the mean of the days of the week for the given interval
If you really wanted a nice imputation, perhaps a lagging average?

```{r impute}
data.avg.per.weekday <- data
data.avg.per.weekday$date <- as.Date(data$date) %>% weekdays()
data.avg.per.weekday <-  data.avg.per.weekday[complete.cases(data.avg.per.weekday), ] %>% group_by(date, interval) %>% summarize(mean=mean(steps, na.rm = TRUE))


data.meanImpute <- data
for (i in 1:nrow(data.meanImpute)){
  if(is.na(data.meanImpute[i, 1])){
    data.meanImpute[i, 1] <- data.avg.per.weekday %>% 
      filter(date == weekdays(as.Date(data[i, 2])) & interval == data[i, 3]) %>%
      .[[3]]
  }
}

print("Checking data:")
any(is.na(data.meanImpute))
# cha-ching
```

Let's use this for a days of the week plot
```{r daysofweek}
# let's get the days of the week reordered properly for the graph
data.meanImpute$date <- weekdays(as.Date(data.meanImpute$date))
data.meanImpute$date <- factor(data.meanImpute$date, levels = 
                                    c("Monday", "Tuesday", "Wednesday", 
                                      "Thursday", "Friday", "Saturday", 
                                      "Sunday"))

data.meanImpute %>% ggplot(aes(x = interval, y = steps, color = date)) +
  geom_line() + theme_bw() + facet_wrap(~date) + 
  labs(title = "Average steps across days of the week\nwith imputed data",
       x = "Time in minutes", y = "Average mean steps")
```

Looks like people are more active saturday night and weekday mornings
